Safe Transferable Regions

### 2016-07-01 paper/introduction.tex
introduction

The use of regions alleviates the performance concern related to
garbage collection described earlier.
However, manual region memory management introduces memory safety
issues such as dangling pointers.
We present a single type system that simultaneously addresses
this safety concern, as well as those described above related to
the use of \emph{transferable data}.

@@

@@

One of the novel contributions of this paper is a type inference algorithm
that eliminates the need for users to write region type annotations.
This, we believe, significantly reduces the impediment to adopt our
approach in practical setting.

@@

\textbf{TO DO:}

As with allocation and deallocation, transferring a memory region is fast, and therefore,
transferring a data structure contained in a region is more efficient
that traversing its objects in heap, and transferring them
independently.
In the \C{SelectVertex} example from
Fig.~\ref{fig:motivating-eg}, the proposed region to contain the
output for each time window $t$ must be transferable due to the
\C{transfer} operation on Line 16. As it is the case with
\C{SelectVertex}, the transferred data is no longer accessed by the
producer, so the transfer operation in our system deallocates the
region once the transfer is complete.

Since transferable regions are first class objects in our setting,
controlling references to and from such regions while ensuring their
safety without significantly diluting the performance advantage of
regions over GC is a challenging exercise.

An important observation, in the context of processes of the kind
described above, is that the data-structures exchanged between them
can be partitioned into sets of fate-sharing objects with common
lifetimes, which makes them good candidates for a region-based memory
management discipline.

>>>>

The use of regions alleviates the performance concern related to
garbage collection described earlier. (It also enables more efficient
serialization/deserialization of data-structures across address spaces,
which is also a significant performance bottleneck in such systems.)
However, manual region memory management introduces memory safety
issues such as dangling pointers.
We present a single type system that simultaneously addresses
this safety concern, as well as those described above related to
the use of \emph{transferable data}.

@@

Formally, our type system introduces \emph{region parameters} (for both
classes and methods) and uses constrained parametric polymorphism
over these parameters, where the constraints capture outlives constraints
between the region parameters.

@@

One of the novel contributions of this paper is a sound type inference algorithm
that eliminates the need for users to write region type annotations.
This, we believe, significantly reduces the impediment to adopt our
approach in practical setting.
Our inference algorithm proceeds in two stages: in the first stage,
we introduce region parameters (for classes and methods), and in
the second stage we infer the outlives constraints between these
parameters (using an interprocedural fixed point computation).
% While the first

@@

%As with allocation and deallocation, transferring a memory region is fast, and therefore,
%transferring a data structure contained in a region is more efficient
%that traversing its objects in heap, and transferring them
%independently.
%In the \C{SelectVertex} example from
%Fig.~\ref{fig:motivating-eg}, the proposed region to contain the
%output for each time window $t$ must be transferable due to the
%\C{transfer} operation on Line 16. As it is the case with
%\C{SelectVertex}, the transferred data is no longer accessed by the
%producer, so the transfer operation in our system deallocates the
%region once the transfer is complete.
%
%Since transferable regions are first class objects in our setting,
%controlling references to and from such regions while ensuring their
%safety without significantly diluting the performance advantage of
%regions over GC is a challenging exercise.
%
%An important observation, in the context of processes of the kind
%described above, is that the data-structures exchanged between them
%can be partitioned into sets of fate-sharing objects with common
%lifetimes, which makes them good candidates for a region-based memory
%management discipline.
